---
title: "Graphics Assignment"
author: "Moses Farley"
format: html
editor: visual
---

## Procedures

All plots are from the Rice river data set. The general procedure for each chart was to select and filter out whatever variables I am analyzing, and to apply a plot to them based off of the "cheat sheet" chart given in the assignment details. Comments are sparse since the first step of this process almost completely mirrors what was done for the function assignment. Charts were generally filled per the examples in given in documentations and from the relevant class slides.

```{r}
#| echo: false
#| warning: false

```

```{r}
#| echo: false
#| label: January rain
rice_data <- get_rice_data()
rice_data %>% #first steps will be filtering & selecting for each unique data used
  select( Rain_in, Month, Weekday) %>%
  filter(Month == "Jan") %>%
  group_by(Weekday) %>%
  summarise(sum(Rain_in)) -> jan_rain
rename(jan_rain, Rain_total = "sum(Rain_in)") -> jan_rain
#creating ggplots is largely the same throughout: select our data, assign aesthetics, and then pick a chart, theme, and create our custom labels if needed
jan_bar <- ggplot(jan_rain, aes(x= Weekday, y = Rain_total)) + geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Rain Inches by Weekday in January", y = "Total Inches")
jan_bar  

```

```{r}
#| echo: false
#| label: January rain (old)
plot.default(jan_rain$Weekday, jan_rain$Rain_total, main = "Rain Inches by Weekday in January (old)", xaxt = 'none',
     xlab = "Weekday", ylab = "Rain inches") #use default to fix points and xaxt to rem old x ticks
axis(side = 1, at = 1:7, labels = jan_rain$Weekday) #new x axis labels

```

```{r}
#| echo: false
#| label: Feb Temps

rice_data %>%
  select(Month, Weekday, AirTempF) %>%
  filter(Month == "Feb") %>%
  group_by(Weekday) %>%
  summarise(mean(AirTempF)) -> temp_feb
rename(temp_feb, Temp = "mean(AirTempF)") -> temp_feb
febTemp_chart <- ggplot(temp_feb, aes(x= Weekday, y = Temp, group = 1)) + geom_line(colour = "aquamarine") + geom_point() + labs(title = 'Average Daily Temps in February') + theme_bw()
febTemp_chart

```

```{r}
#| echo: false
#| title: Feb Temps (bad chart)

plot.default(temp_feb$Weekday, temp_feb$Temp, type = "p", main = "Average Daily Temps in February (old)", xaxt = 'none',
     xlab = "Weekday", ylab = "Avg Temp F") #use default to fix points and xaxt to rem old x ticks
axis(side = 1, at = 1:7, labels = temp_feb$Weekday)
```

```{r}
#| echo: false
#| label: valentines day

rice_data %>%
  select(Day, Month, Depth_ft) %>%
  filter(Month == "Feb" & Day == 14) %>%
  select(Depth_ft) -> water_depth
val_dep <- ggplot(water_depth, aes(x = row_number(Depth_ft), y = Depth_ft)) + geom_col(colour = "darkcyan") + labs(title = 'Water Depth on Valentines Day', x = 'Observations over day', y = 'Depth (ft)') + theme_classic()
val_dep
```

```{r}
#| echo: false
#| title: Valentines day (old)

plot(row_number(water_depth), water_depth$Depth_ft, main = "Water Depth on Valentines Day (old)", xlab = "Observations over day", ylab = "Water Depth (ft)")

```

```{r}
#| echo: false
#| label: Total Rain and Median Temps

rice_data %>%
  select(Month, Rain_in, AirTempF) %>%
  group_by(Month) %>%
  summarise(sum(Rain_in), median(AirTempF)) -> mth_chart
rain_temps_chart <- ggplot(mth_chart, aes(x = `sum(Rain_in)`, y = `median(AirTempF)`)) +
  geom_point(colour = "red") + geom_text(aes(label = Month), vjust = -0.5) + labs(title = 'Monthly Air Temp and Rainfall', x = 'Total Rainfall Inches', y = 'Median AirTemp F') + theme_dark()
rain_temps_chart
```

```{r}
#| echo: false
#| label: Total Rain and Median Temps (old)
plot(mth_chart$`sum(Rain_in)`, mth_chart$`median(AirTempF)`, main = "Monthly Air Temp and Rainfall (old)", xlab = 'Total Rainfall Inches', ylab = 'Median AirTemp F')

```

```{r}
#| echo: false
#| label: Jan 8th week wind speed
rice_data %>%
  select(Month, Day, WindDir, Weekday) %>%
  filter(Month == 'Jan') %>%
  filter(Day >= 5 & Day <= 11) -> tep
wind_chart <- ggplot(tep, aes(x = Weekday, y = WindDir)) + geom_boxplot(colour = "darkseagreen") + labs(title = "Wind Direction Jan 8th Week", y = "Wind Direction (degrees)")
wind_chart
```

```{r}
#| echo: false
#| label: Jan 8th week wind speed (old)

plot(tep$Weekday, tep$WindDir, main = "Wind Direction Jan 8th Week", ylab = "Wind Direction (degrees)", xlab = "Weekday")
```
