---
title: "Data Types In Class Activity"
author: "YOUR NAME HERE"
format: 
  html:
    code-tools: true
    toc: false
    exec: true
    echo: false
---

## Overview & Objective

This is the first foray into how to actually use `R` and focuses on understanding some of the most basic data types.

## Student Learning Objectives

At the end of this topic, you should be able to:

-   Describe characteristics of numeric, character, logical, and date data types.\
-   Use introspection and coercion for basic data types.
-   Perform operations on included data types.

## Activity

The following markdown has a set of questions followed by a chunk for you to put in code to answer the question. Fill them all out as indicated.

### Numeric Data

What is your favorite number

```{r}
myFavNum <- 60
myFavNum
```

The ancient Egyptians were quite cleaver at engineering and building things. The [Rhind Papyrus](https://en.wikipedia.org/wiki/Rhind_Mathematical_Papyrus), which dates to 1650 BCE, indicates that they estimated the ratio between the circumfrence and the diameter of a circle to be 256 divided by 81. Mathematically is this more or less than what we estimate $\pi$ as? What is the magnitude of the difference?

```{r}
pi
ePi <- 256/81
ePi_error <- ePi - pi
ePi_error
```

The function `sqrt()` provides you the square root of a number. Assign a variable named `root2` to hold the value of `sqrt(2)`.

```{r}
root2 <- sqrt(2)
root2
```

What is the value contained in the variable below named `sqrOfNegative2`? Explain the result and make any modifications necessary to be more explicit.

```{r}
sqrOfNegative2 <- (-2)^2
sqrOfNegative2
```

### Character Data

Create variables for your first and last names (and give them proper variable names).

```{r}
firstName <- "Moses"
lastName <- "Farley"

```

Use the functions `print()` and `cat()` to display these values. Are these two functions identical?

```{r}
print(firstName)
cat(lastName)
```

Create a new variable, `properName`, which is your last name, a commma and space, followed by your first name. Use the variables above instead of writing out the letters for your names.

```{r}
properName = paste0(lastName, firstName)
cat(lastName, firstName, sep = " ")
properName[1]
```

The function `nchar()` returns the number of characters in a character variable. Use it on your `properName`.

```{r}
nchar(properName)
```

### Logical Data

Create the variables `trueResult` and `falseResult` and assign them the proper logical values for R.

```{r}
trueResult <- TRUE
falseResult <- FALSE
```

Using the variables `trueResult` and `falseResult` demonstrate logical operators for equality, inequality, and negation.

```{r}
trueResult == falseResult
trueResult != falseResult
!trueResult
!falseResult
```

Using the variables `trueResult` and `falseResult` demonstrate AND operator in R for all pairwise combinations (e.g, for the and operator, show how to evaluate `trueResult` and `trueResult`, `trueResult` and `falseResult`, `falseResult` and `trueResult`, and `falseResult` and `falseResult`)

```{r}
trueResult & trueResult
trueResult & falseResult
falseResult & trueResult
falseResult & falseResult
```

Do the same as the previous question for the approriate OR operator.

```{r}
trueResult | trueResult
trueResult | falseResult
falseResult | trueResult
falseResult | falseResult
```

In the example above for the square root, the variable `root2` has the square root of the number $2$. Test the equality (e.g., use the `==` operator) of this varible times itself to the value $2$. Discuss this with your neighbor. Does this make sense?

```{r}
root2^2 == 2
```

### Dates

When did you first become a college student? Create a variable representing this date.

```{r}
library(lubridate)
dateStart <- mdy("09-01-2016")
```

How many days was it between then and now?

```{r}
today() - dateStart
```
